```{r knitsetup, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
source('../../global_options.R')
opts_knit$set(base.dir='./', out.format='md')
opts_chunk$set(prompt=FALSE, comment='', results='markup', cache = cch, cache.path = '../../book_cache/', fig.path = '../../book_figures/', fig.align = 'center')


knit_hooks$set(plot = function(x, options) {
  paste('<figure><img src="',
        opts_knit$get('base.url'), paste(x, collapse = '.'),
        '"><figcaption>', options$fig.cap, '</figcaption></figure>',
        sep = '')
})

fn = local({
  i = 0
  function(x) {
    i <<- i + 1
    paste('Figure ', i, ': ', x, sep = '')
  }
})
# See yihui.name/knitr/options for more Knitr options.
##### Put other setup R code here
clrs_hcl <- function(n) {
  hcl(h = seq(230, 0, length.out = n), 
      c = 60, l = seq(10, 90, length.out = n), 
      fixup = TRUE)
  }

dmod <- lm(price ~ cut, data = diamonds)
cuts <- data.frame(cut = unique(diamonds$cut), 
                   predict(dmod, data.frame(cut = unique(diamonds$cut)), 
                           se = TRUE)[c("fit", "se.fit")])
# end setup chunk
```

### 2.8. Plotting error bars (ggplot2 way)

As mentioned above, when plotting error bars ```ggplot2``` is much easier. Whether this is because of the genereal ongoing discussion about the usefulness of these plots I do not want to judge.

Anyway, plotting error bars in ```gglplot2``` is as easy as everything else...

```{r gg err, tidy = FALSE, fig.cap = fn("a basic dotplot with error bars produced with ggplot2"), fig.align = 'center'}
errbar_ggplot <- ggplot(cuts, aes(cut, fit, ymin = fit - se.fit, 
                                  ymax=fit + se.fit))
g_err <- errbar_ggplot + 
  geom_pointrange() +
  coord_flip() +
  theme_classic()

print(g_err)
```

Especially, when plotting them as part of a bar plot.

```{r gg err bar, tidy = FALSE, fig.cap = fn("a ggplot2 bar plot with error bars"), fig.align = 'center'}
g_err <- errbar_ggplot + 
  geom_bar(fill = "grey80") +
  geom_errorbar(width = 0.2) +
  coord_flip() +
  theme_classic()

print(g_err)
```

Just as before (with the box wdths) though, applying this to each panel is a little more complicated...

```{r gg facet err bar, tidy = FALSE, concordance = TRUE, fig.cap = fn("a ggplot2 panel bar plot with error bars and modified fill colours"), fig.align = 'center'}
errbar_ggplot_facets <- ggplot(diamonds, aes(x = color, y = price))

### function to calculate the standard error of the mean
se <- function(x) sd(x)/sqrt(length(x))

### function to be applied to each panel/facet
myFun <- function(x) {
  data.frame(ymin = mean(x) - se(x), 
             ymax = mean(x) + se(x), 
             y = mean(x))
  }

g_err_f <- errbar_ggplot_facets + 
  stat_summary(fun.y = mean, geom = "bar", 
               fill = clrs_hcl(7)) + 
  stat_summary(fun.data = myFun, geom = "linerange") + 
  facet_wrap(~ cut) +
  theme_bw()

print(g_err_f)
```

Still, trust me on this, much much easier than to achieve the same result in ```lattice```.
