```{r knitsetup, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
source('../../global_options.R')
opts_knit$set(base.dir='./', out.format='md')
opts_chunk$set(prompt=FALSE, comment='', results='markup', cache = cch, cache.path = '../../book_cache/', fig.path = '../../book_figures/', fig.align = 'center')


knit_hooks$set(plot = function(x, options) {
  paste('<figure><img src="',
        opts_knit$get('base.url'), paste(x, collapse = '.'),
        '"><figcaption>', options$fig.cap, '</figcaption></figure>',
        sep = '')
})

fn = local({
  i = 0
  function(x) {
    i <<- i + 1
    paste('Figure ', i, ': ', x, sep = '')
  }
})
# See yihui.name/knitr/options for more Knitr options.
##### Put other setup R code here
clrs_hcl <- function(n) {
  hcl(h = seq(230, 0, length.out = n), 
      c = 60, l = seq(10, 90, length.out = n), 
      fixup = TRUE)
  }

# end setup chunk
```

### 2.7. Histograms and densityplots (ggplot2 way)

Much like with the box and whicker plot, the default settings of ```ggplot2``` are quite a bit nicer for both histogram and densityplot.

```{r gg hist, tidy = FALSE, fig.cap = fn("a basic histogram produced with ggplot2"), fig.align = 'center'}
hist_ggplot <- ggplot(diamonds, aes(x = price))

g_his <- hist_ggplot +
  geom_histogram()

print(g_his)
```

One thing that is really nice about the ```ggplot2``` densityplots that it is so easy to fill the area under the curve which really helps the visual represeantation of the data.

```{r gg dens, tidy = FALSE, fig.cap = fn("a basic density plot produced with ggplot2"), fig.align = 'center'}
dens_ggplot <- ggplot(diamonds, aes(x = price))

g_den <- dens_ggplot +
  geom_density(fill = "black", alpha = 0.5)

print(g_den)
```

Just as before, we are encountering again the rather peculiar way of ```ggplot2``` to adjust certain default settings to suit our needs (likes). In we wanted to show percentages instead of counts for the histograms, we again need to use the strange ```..something..``` syntax.

Anothr thing I want to highlight in the following code chunck is the way to achive binary conditioning in ```ggplot2```. This is done as follows:

```
facet_grid(g ~ f)
```

where, again, ```g``` and ```f``` are the two varibles used for the conditioning.

```{r gg facet hist, tidy = FALSE, fig.cap = fn("a ggplot2 panel histogram with y-axis labelled according to percentages"), fig.align = 'center', message = FALSE}
g_his <- hist_ggplot +
  geom_histogram(aes(y = ..ncount..)) +
  scale_y_continuous(labels = percent_format()) +
  facet_grid(color ~ cut) + 
  ylab("Percent")

print(g_his)
```

Similar to our ```lattice``` approach we're going to rotate the x-axis labels by 45 degrees.

```{r gg facet density, tidy = FALSE, fig.cap = fn("a ggplot2 panel density plot conditioned according to 2 variables"), fig.align = 'center'}
dens.ggplot <- ggplot(diamonds, aes(x = price))

g_den <- dens_ggplot +
  geom_density(fill = "black", alpha = 0.5) +
  facet_grid(color ~ cut) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(g_den)
```

Ok, antoher thing we might want to show is the value of a certain estimated value (like the mean of our sample) including error bars.
